### 2014-08-18 doc/thesis/typedlua.tex
Adding text about description files

>>>>

\section{Description files}

The Typed Lua compiler also handles description files.
The programmer can use description files for two purposes:
export a statically typed interface for an existing Lua
module, without coverting it to Typed Lua,
or export a statically typed interface for a Lua module
that is written in C.

\begin{figure}[!ht]
\textbf{The complete syntax of description files}\\
\dstart
\begin{align*}
\textit{description} & ::= \; \textit{desclist}\\
\textit{desclist} & ::= \; \textit{descitem} \; \{ \textit{descitem} \}\\
\textit{descitem} & ::= \; \textit{typedid} \; | \;
  \textit{interface} \; | \;
  \textit{userdata}\\
\textit{typedid} & ::= \; \textit{id} \; \texttt{`:'} \; \textit{type}\\ 
\textit{interface} & ::= \; \textbf{interface} \; \textit{typedec}\\
\textit{userdata} & ::= \; \textbf{userdata} \; \textit{typedec}\\
\textit{typedec} & ::= \; \textit{Name} \; { \; \textit{decitem} \; } \; \textbf{end}\\
\textit{decitem} & ::= \; \textit{idlist} \; \texttt{`:'} \; \textit{idtype}\\
\textit{idtype} & ::= \; \textit{type} \; | \; \textit{methodtype}\\
\textit{idlist} & ::= \; \textit{id} \; \{ \; \texttt{`,'} \; \textit{id} \; \}\\
\textit{id} & ::= \; [\textbf{const}] \; \textit{Name}
\end{align*}
\dend
\caption{The concrete syntax of Typed Lua description files}
\label{fig:tldsyntax}
\end{figure}

The Typed Lua compiler searches for description files when it cannot
not find the respective Typed Lua file that is the argument of
\texttt{require}, as we can also use description files to give and
export the type of modules.

The following example shows the description file for \texttt{lmd5},
a MD5 digest library for Lua that is written in C:
\begin{verbatim}
    userdata md5_context
      __tostring : (self) -> (string)
      clone : (self) -> (self)
      digest : (self|string, value?) -> (string)
      new : () -> (self)
      reset : (self) -> (self)
      update : (self, string*) -> (self)
      version : string
    end

    __tostring : (md5_context) -> (string)
    clone : (md5_context) -> (md5_context)
    digest : (md5_context|string, value) -> (string)
    new : () -> (md5_context)
    reset : (md5_context) -> (md5_context)
    update : (md5_context, string*) -> (md5_context)
    version : string
\end{verbatim}

This description file exports the type \texttt{md5\string_context}
through the \texttt{userdata} declaration, and the table type
that represents the module.
Now we can use the Typed Lua compiler to check for type errors on
our use of the library:
\begin{verbatim}
    local m = require "md5"
    local x = m.new()
    local y = x:clone()
    local z = m.clone("foo")             -- compile-time error
    print(x:digest() == m.digest(x))
\end{verbatim}

In this example there is a type error because function \texttt{clone}
expects a value of type \texttt{md5\string_context} instead of a value
of type \texttt{string}.

The description files are the mechanism that we used to include the
typing of the Lua standard libraries inside Typed Lua.
